---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CellDiv

<!-- badges: start -->
<!-- badges: end -->

The goal of CellDiv is to calculate the Shannon Diversity Index for the cell types of interest to the user and to allow the identification of the combination of cell types that best distinguish dichotomous outcomes

## Installation

You can install the development version of CellDiv from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("WenqinXie/CellDiv")
```

## Usage

Generate sample data
```{r}
#cdata
set.seed(1)
cdata1 <- matrix(sample(1000:5000, 500, replace = TRUE), 50, 10)
row.names(cdata1) <- paste0("A",1:50)
colnames(cdata1) <- c(paste0("CellType", 1:10))
cdata1 <- as.data.frame(cdata1)

set.seed(1)
cdata2 <- matrix(sample(2000:4000, 400, replace = TRUE), 40, 10)
row.names(cdata2) <- paste0("A",1:40)
colnames(cdata2) <- c(paste0("CellType", 1:10))
cdata2 <- as.data.frame(cdata2)

cdata <- rbind(cdata1, cdata2)
print(cdata[1:5, 1:5])
```

```{r}
#metadata
set.seed(1)
metadata <- data.frame(sample = row.names(cdata),
condition = c(rep("SD/PD", 50), rep("CR/PR", 40)))
metadata$condition <- as.factor(metadata$condition)
print(head(metadata))
```

Calculate diversity score for each sample
```{r}
library(CellDiv)
divscores <- celltdiv(cdata = cdata, cellsets = colnames(cdata)[1:8])
print(head(divscores))
```

Average AUC values are calculated after multiple samples for all combinations of cell types of interest
```{r}
cdiv <- commroc(cdata = cdata, metadata = metadata, 
        cellsets = colnames(cdata)[1:8], sami = 500, ncores = 10)
print(head(cdiv))
```

The combination of cell types corresponding to the maximum AUC value was selected
```{r}
cellsets <- maxroc(acroc = cdiv)
print(cellsets)
```
Plot relative percentage of cell types
```{r}
#"SD/PD"
plotProp(cdata = cdata1, 
         slcells = strsplit(cellsets$cellsets, "; ")[[1]], 
         xtitle = "No_Response",
axistext.x = FALSE, bar_width = 0.5, flow = TRUE, geomtext = FALSE,
color = c("#00468BFF", "#925E9FFF", "#AD002AFF", "#FDAF91FF", "#AFFF1F36", "#FDAF352F"))

#"CR/PR"
plotProp(cdata = cdata2, 
         slcells = strsplit(cellsets$cellsets, "; ")[[1]], 
         xtitle = "Response",
axistext.x = FALSE, bar_width = 0.5, flow = TRUE, geomtext = FALSE,
color = c("#00468BFF", "#925E9FFF", "#AD002AFF", "#FDAF91FF", "#AFFF1F36", "#FDAF352F"))
```

Plot the diversity score comparison between the two conditions
```{r}
data <- merge(divscores, metadata, by = "sample", sort = FALSE)
plotCC(data = data, x = "condition", y = "divscore",
xlabs = "Condition", ylabs = "DiversityScore",
color = c("#00AED7", "#FD9347"),
label.x = 1.5, label.y = 2.1,
comparisons = NULL)
```

